---
title: "P8105 Homework 1"
output: github_document
date: "September 24th, 2022"
---

```{r, include = FALSE}
library(tidyverse)
```

## Problem 1: Exploratory Data Analysis

In this homework, we will be using the `penguins` dataset from `palmerpenguins` package. 
```{r}
data("penguins", package = "palmerpenguins")
```
We can use the `head` function to take a look at the first 5 rows of the data and to see what variables are present. Using `nrow` and `ncol`, we can take a look at how many observations and variables are in our dataset.
```{r}
head(penguins, 5)
nrow(penguins)
ncol(penguins)
```

We can see that the `penguins` dataset has 8 variables and 344 observations. For each penguin, it contains information about the species, sex, island, year they were recorded, as well as several size measurements, including bill length and depth, flipper length, and body mass.

The mean flipper length in the data is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` millimeters.

To do further data exploration, we want to graph the relationship between flipper length and bill length, by penguin species. We can do so with a scatterplot using `ggplot`.

```{r plot, echo = FALSE, warning = FALSE}
ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + 
  geom_point() + 
  labs(x = "Bill Length (mm)", y = "Flipper Length (mm)")

ggsave("scatter_plot.pdf", height = 4, width = 6)

```

## Problem 2: Working with dataframes

We will now create a dataframe using `tibble` containing:

* a random sample of size 10 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 10
* a factor vector of length 10, with 3 different factor “levels”

```{r dataframe}
p2_df = tibble(
  v_numeric = rnorm(10),
  v_logical = v_numeric > 0,
  v_char = c("Apple", "Milk", "Kiwi", "Cucumber", "Soda", 
             "Pepper", "Strawberry", "Mango", "Spinach", "Coffee"),
  v_factor = factor(c("Fruit", "Beverage", "Fruit", "Vegetable", 
                      "Beverage", "Vegetable", "Fruit", "Fruit", "Vegetable", "Beverage"))
)

head(p2_df)
```
Next, we will try to take the mean of each variable in our `p2_df` dataframe.

```{r mean}
mean(pull(p2_df, var = 1))
mean(pull(p2_df, var = 2))
mean(pull(p2_df, var = 3))
mean(pull(p2_df, var = 4))
```

When we try to calculate the mean of each variable in `p2_df`, we get the mean values for `v_numeric` and `v_logical`, while getting an output of `NA` for the `v_char` and `v_factor`. Therefore, the mean function appears to only work for numeric or logical variables.

Next, we will try to convert the logical, character, and factor variables into numeric variables using `as.numeric  `. 

```{r, results = "hide", warning = FALSE}
p2_df$v_logical = as.numeric(p2_df$v_logical)
p2_df$v_char = as.numeric(p2_df$v_char)
p2_df$v_factor = as.numeric(p2_df$v_factor)
 
mean(pull(p2_df, var = 2))
mean(pull(p2_df, var = 3))
mean(pull(p2_df, var = 4))
```

After using `as.numeric`, we can see the `v_logical` replaces `TRUE` and `FALSE` with a string of 1's and 0's. We get a mean of `r mean(pull(p2_df, var = 2))`, which is the same as what we calculated as a logical variable, indicating that R takes the average number of successes observed in our dataset.

Similarly `as.numeric` assigning a numeric value to each level of our factor variable, `v_factor`. Our mean is `r mean(pull(p2_df, var = 4))`. This allowed R to calculate a mean from these now numeric values. This, however, may not be meaningful, as we had a nominal factor variable.

Finally, using `as.numeric` on the character variable did not assign numeric variables to the structure, and instead, coerced all values to `NA`. This indicates that character variables with no structure or order cannot be converted to numeric values. As such, we could not calculate the mean for `v_char`.
