---
title: "P8105 Homework 1"
output: github_document
date: "2022-09-21"
---

```{r, include = FALSE}
library(tidyverse)
```

## Problem 1: Exploratory Data Analysis

In this homework, we will be using the `penguins` dataset from `palmerpenguins` package. 
```{r}
data("penguins", package = "palmerpenguins")
```
We can use the `head` function to take a look at the first 5 rows of the data and to see what variables are present. Using `nrow` and `ncol`, we can take a look at how many observations and variables are in our dataset
```{r}
head(penguins, 5)
nrow(penguins)
ncol(penguins)
```

We can see that the `penguins` dataset has 8 variables and 344 observations. For each penguin, it contains information about the species, sex, island, year they were recorded, as well as several size measurements, including bill length and depth, flipper length, and body mass.

The mean flipper length in the data is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` millimeters.

To do further data exploration, we want to graph the relationship between flipper length and bill length, by penguin species. We can do so with a scatterplot using `ggplot`.

```{r plot, echo = FALSE, warning = FALSE}
ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + 
  geom_point() + 
  labs(x = "Bill Length (mm)", y = "Flipper Length (mm)")

ggsave("scatter_plot.pdf", height = 4, width = 6)

```

## Problem 2


